---
title: "Homework 6"
author: "Maxwell Gress"
editor: visual
format:
  pdf:
    embed-resources: true
execute:
  echo: true
  message: false
  warning: false
---

```{r setup, echo = FALSE}
library(tidyverse)
library(tidytext)
library(schrute)
data("theoffice")
```

# The Office

Let's continue to work with *The Office* text from the [schrute](https://cran.r-project.org/web/packages/schrute/index.html) package.

# Sentiment Analysis

### Over Time

Calculate the sentiment score (positive - negative) for each episode, and plot this variable over time.

```{r}
#Creating a dataset with an overall episode number (instead of season + episode)
theoffice0 <- theoffice |> 
  arrange(season, episode) |>
  #consequtive_id creates a counter everytime the combo of season and episode changes
  mutate(overall_episode = consecutive_id(season, episode))
```

* Jerry: Could try "theoffice |> mutate(overall_episode = paste0(season, ".", episode" if you want to see which episode is from which season.

```{r}
#Creating dataset with each line and the episode associated with that line
theoffice_textdf <- theoffice0 |> 
  select(overall_episode, text) #reordering
```

```{r}
#Creating a list of words and episode they are associated with
theoffice_words <- theoffice_textdf |> 
  unnest_tokens(word, text) #tokenize
```

```{r}
theoffice_words |> 
  #Joining sentiments from tidytext with the words in the office dataset
  inner_join(sentiments, by = "word") |>
  #Grouping by overall episode
  group_by(overall_episode) |> 
  #Creating sentiment score for each episode
  summarize(pos_score = sum(sentiment == "positive"),
            neg_score = sum(sentiment == "negative"),
            sent_score = pos_score - neg_score) |> 
  #Plotting sentiment score over overall episodes
  ggplot(aes(x = overall_episode, y = sent_score)) +
  geom_line() +
  labs(x = "Overall Episode", y = "Sentiment Score",
       title = "Sentiment Score of The Office Episodes Over Time") +
  theme_minimal()
```

* Jerry: You could also try making another sentiment score that lists only the seasons in the x-axis, while still keeping the score based on episodes per season.

### By Rating

Create a scatterplot of the `imdb_rating` versus the sentiment score (positive - negative), where the case is again an episode. Decide whether a linear model is a good fit to this relationship (optional, if you have had modeling classes, what model looks best?). Comment on what you find.

```{r}
theoffice_words |> 
  inner_join(sentiments, by = "word") |>
  group_by(overall_episode) |> 
  summarize(pos_score = sum(sentiment == "positive"),
            neg_score = sum(sentiment == "negative"),
            sent_score = pos_score - neg_score) |> 
  #Joining IMBD Rating back into the new data and sentiment
  left_join(theoffice0 |> 
              distinct(overall_episode, imdb_rating), by = "overall_episode") |> 
  ggplot(aes(x = imdb_rating, y = sent_score)) +
  geom_point() +
  labs(x = "IMDB Rating", y = "Sentiment Score",
       title = "Sentiment Score vs IMDB Rating for The Office Episodes") +
  theme_minimal() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") #Adding a linear model fit
```

The linear relationship is positive but not very strong. The points are widely scattered around it, and there seems to be high variability in sentiment scores at every IMBD rating level. However, we can say that higher IMBD ratings are associated with higher sentiment scores, though we can't say that with much confidence.

* Jerry: You can try to color each observation based on season. You would have to mutate season before coloring the ggplot.
* "theoffice |> mutate(seasonv2 = factor(season))"

### Sentiment Negation

Identify at least a few negations that change the sentiment values based on the text. Use this to recreate the graphs above (or do this first, then create the graphs).

```{r, eval = FALSE}
#Identifying negation words using bigrams
theoffice_textdf |>
  unnest_tokens(bigram, text, token = "ngrams", n = 2) |>
  drop_na(bigram) |>
  separate(bigram,
           into = c("word1", "word2"),
           sep = " ",
           remove = FALSE) |> 
  inner_join(sentiments,
             by = c("word2" = "word")) |>
  count(word1, sort = TRUE)
  
```

The word "not" counted 472 times. The word "don't" counted 253 times. We could consider this a negation word. "No" counted 236 times. "Doesn't" counted 49 times. "Didn't" counted 48 times. "Can't" counted 47 times. "Never" counted 37 times. "Isn't" counted 17 times. "Wasn't" counted 17 times. I am going to use this list of words as my negation words for The Office.

```{r}
reversal_words <- c("not", "don't", "no", "doesn't",
                    "didn't", "can't", "never", "isn't", "wasn't")
```

```{r}
theoffice_textdf |>
  unnest_tokens(bigram, text, token = "ngrams", n = 2) |>
  separate(bigram,
           into = c("word1","word2"),
           sep = "\\s+") |>
  inner_join(sentiments |> rename(sentiment0 = sentiment),
             by = c("word2" = "word")) |>
  #Reversing sentiment if the first word is a reversal word
  mutate(sentiment = case_when(
        word1 %in% reversal_words & sentiment0 == "positive" ~ "negative",
        word1 %in% reversal_words & sentiment0 == "negative" ~ "positive",
        TRUE ~ sentiment0)) |>
  group_by(overall_episode) |>
  summarize(
    pos_score = sum(sentiment == "positive"),
    neg_score = sum(sentiment == "negative"),
    sent_score = pos_score - neg_score,
    pos_score0 = sum(sentiment0 == "positive"),
    neg_score0 = sum(sentiment0 == "negative"),
    sent_score0 = pos_score0 - neg_score0,
    .groups = "drop") |> 
  ggplot(aes(x = overall_episode, y = sent_score)) +
  geom_line() +
  labs(x = "Overall Episode", y = "Sentiment Score",
       title = "Sentiment Score of The Office Episodes Over Time") +
  theme_minimal()
```

```{r}
theoffice_textdf |>
  unnest_tokens(bigram, text, token = "ngrams", n = 2) |>
  separate(bigram,
           into = c("word1","word2"),
           sep = "\\s+") |>
  inner_join(sentiments |> rename(sentiment0 = sentiment),
             by = c("word2" = "word")) |>
  #Reversing sentiment if the first word is a reversal word
  mutate(sentiment = case_when(
    word1 %in% reversal_words & sentiment0 == "positive" ~ "negative",
    word1 %in% reversal_words & sentiment0 == "negative" ~ "positive",
    TRUE ~ sentiment0)) |>
  group_by(overall_episode) |>
  summarize(
    pos_score = sum(sentiment == "positive"),
    neg_score = sum(sentiment == "negative"),
    sent_score = pos_score - neg_score,
    pos_score0 = sum(sentiment0 == "positive"),
    neg_score0 = sum(sentiment0 == "negative"),
    sent_score0 = pos_score0 - neg_score0,
    .groups = "drop") |> 
  left_join(theoffice0 |> 
              distinct(overall_episode, imdb_rating), by = "overall_episode") |> 
  ggplot(aes(x = imdb_rating, y = sent_score)) +
  geom_point() +
  labs(x = "IMDB Rating", y = "Sentiment Score",
       title = "Sentiment Score vs IMDB Rating for The Office Episodes") +
  theme_minimal() +
  geom_smooth(method = "lm", se = FALSE, color = "blue")
```

The graphs don't seem to change much to me.
